---
  title: "Skin Metabolic Activity"
author: "Don R Schoolmaster Jr. and Jay T Lennon"
date: "`r format(Sys.time(), '%d %B, %Y')`"
header-includes:
  - \usepackage{array}
output: pdf_document
geometry: margin=2.54cm
---

# Set working directory
```{r}
rm(list=ls())
getwd()
setwd("~/GitHub/skin")
```

# Load packages
```{r}
require("MASS")
require("psych")
```

# Read in data 
```{r}
ages.raw.count <- read.table("data/skin.age.counts.txt", sep = "\t", header = T)

# Remove subject "G1"
ages <- ages.raw.count[ ! ages.raw.count$subject == "G1", ]
```

# Create dummy variables for multinomial variables
```{r}
# Shower frequency
  # "1" = < three times per week
  # "2" = every other day
  # "3" = once per day
show.f2 <- ifelse(ages$show.freq==2,1,0)
show.f3 <- ifelse(ages$show.freq==3,1,0)

# Shower length
  # "1" = 0 - 5 mins
  # "2" = 5 - 15 mins
  # "3" = > 15 mins
show.l2 <- ifelse(ages$show.leng==2,1,0)
show.l3 <- ifelse(ages$show.leng==3,1,0)

# Last shower
  # "1" = > 48 hrs
  # "2" = 24 - 48 hrs
  # "3" = < 24 hrs
show.last2 <- ifelse(ages$last.show==2,1,0)
show.last3 <- ifelse(ages$last.show==3,1,0)

# Merge dummry variables into dataframe
ages.2 <- data.frame(ages[,c(1:6,11:16)], show.f2,show.f3,show.l2, 
          show.l3,show.last2,show.last3)

# Set up a design matrix that we might use later
X <- data.frame(ages.2[,-c(1,3:5)])


summary(f.act<-glm.nb(ages$active~.+offset(log(X$total)),data=X[,-2]))
who<-names(coef(f.act))[summary(f.act)$coef[,4]<0.05]
who<-who[-1]
summary(f.act.int<-glm.nb(ages$active~age*show.l2+age*show.f3+offset(log(X$total)),data=X[,who]))
cf<-coef(f.act.int)
summary(f.act.int<-glm.nb(ages$active~age*show.l2+offset(log(X$total)),data=X[,who]))
cf<-coef(f.act.int)
cf
plot(log(ages$active/ages$total)~X$age,pch=21,bg=X$show.l2+1)
curve((cf[1]+(cf[2])*x),from=-40,to=40,add = T)
curve((cf[1]+cf[3]+(cf[2]+cf[5])*x),from=-40,to=40,add = T)
