---
  title: "Skin Metabolic Activity"
author: "Don R Schoolmaster Jr. and Jay T Lennon"
date: "`r format(Sys.time(), '%d %B, %Y')`"
header-includes:
  - \usepackage{array}
output: pdf_document
geometry: margin=2.54cm
---

# Set working directory
```{r}
rm(list=ls())
getwd()
setwd("~/GitHub/skin")
```

# Load packages
```{r}
require("MASS")
require("psych")
```

# Read in data 
```{r}
ages.raw.count <- read.table("data/skin.age.counts.txt", sep = "\t", header = T)

# Remove subject "G1"
ages <- ages.raw.count[ ! ages.raw.count$subject == "G1", ]
```

# Create dummy variables for multinomial variables
```{r}
# Shower frequency
  # "1" = < three times per week
  # "2" = every other day
  # "3" = once per day
show.f2 <- ifelse(ages$show.freq==2,1,0)
show.f3 <- ifelse(ages$show.freq==3,1,0)

# Shower length
  # "1" = 0 - 5 mins
  # "2" = 5 - 15 mins
  # "3" = > 15 mins
show.l2 <- ifelse(ages$show.leng==2,1,0)
show.l3 <- ifelse(ages$show.leng==3,1,0)

# Last shower
  # "1" = > 48 hrs
  # "2" = 24 - 48 hrs
  # "3" = < 24 hrs
show.last2 <- ifelse(ages$last.show==2,1,0)
show.last3 <- ifelse(ages$last.show==3,1,0)

# Merge dummry variables into dataframe
ages.2 <- data.frame(ages[,c(1:6,11:16)], show.f2,show.f3,show.l2, 
          show.l3,show.last2,show.last3)

# Set up a design matrix that we might use later
X <- data.frame(ages.2[,-c(1,3:5)])

# Center age for later interpretation
cat("average age in survey is", round(mean(ages$age),1))
X$age <- X$age-mean(X$age)
```

# Perform SEM: parameter estimates, errors, and significance
```{r}
# Active
summary(f.act<-glm.nb(ages$active~.+offset(log(X$total)),data=X[,-2]))

# Dormant
summary(f.dor<-glm.nb(ages$dormant~.+offset(log(X$total)),data=X[,-2]))

# Dead
summary(f.ded<-glm.nb(ages$dead~.+offset(log(X$total)),data=X[,-2]))

# Total
summary(f.tot<-glm.nb(ages$total~.,data=X[,-2],maxit=1000))
```

# Which variables significantly affect metabolic classes?
```{r}
# Active
act.n <- rownames(summary(f.act)$coef)[which(summary(f.act)$coef[,4]<0.05)] 
act.p <- summary(f.act)$coef[which(summary(f.act)$coef[,4]<0.05)] 
t(data.frame(act.n, act.p))
# Finding: active cells decrease with age; increase with shower freq (f3) and length (l2)
# Finding: marginal (p = 0.085) positive effect of tanning on active bacteria

# Dormant
dor.n <- rownames(summary(f.dor)$coef)[which(summary(f.dor)$coef[,4]<0.05)]
dor.p <-summary(f.dor)$coef[which(summary(f.dor)$coef[,4]<0.05)]
t(data.frame(dor.n,dor.p))
# Finding: dormant cells increase with increasing age and shower length (l2 and l3)

# Dead
ded.n <- rownames(summary(f.ded)$coef)[which(summary(f.ded)$coef[,4]<0.05)] 
ded.p <-summary(f.ded)$coef[which(summary(f.ded)$coef[,4]<0.05)]
t(data.frame(ded.n,ded.p))
# Finding: dead cells decrease with increasing shower length (l3)

# Total
tot.n <- rownames(summary(f.tot)$coef)[which(summary(f.tot)$coef[,4]<0.05)] 
tot.p <- summary(f.tot)$coef[which(summary(f.tot)$coef[,4]<0.05)] # parms
t(data.frame(tot.n,tot.p))
# Finding: total cells decrease with age and shower length (l2 and l3)
# Finding: total cells increase with ecze

# age and shower length both reduce cells and affect distribution of cells
# shower freq increases active cells
# eczema increases total but does not affect the distribution across metabolic classes
```

# Identify potential dormancy cues: coefs of factors that are of one sign for active and dead cells and opposite sign for dormant cells
```{r}
# Make matrix of coefficiencts
coef.matrix <- rbind(coef(f.act),coef(f.dor),coef(f.ded))
rownames(coef.matrix)<-c('active','dormant','dead')

# Consider variables that were significant above
coef.matrix[,"age"] # dormancy inducing mechanism (and reduces total cell number)
coef.matrix[,"show.f3"] # show.f3 is a dormancy reducing mechanism
coef.matrix[,"show.l3"] #show.l3 is NOT a dormancy mechanism, but 
#works to preferentially remove dead cells--this interp is supported by the 
#significant effects on total cell count see detail below
coef.matrix[,"show.l2"] # show length is NOT a dormancy mechanisms, but
#works to preferentially remove dead cells

# Notice that shower freq and last show should be on average inversely related and the 
# coefs show that they tell the same story shower freq tends to reduce dormancy
# so, the longer its been since showering the higher proportion is in the dormant category
# Don?: is this sort of trivial result? Also, show.last2 and show.last3 are never significant in SEM. 
# Don?: perhaps remove show.last from dataset, or just leave it in and don't bother highlighting
```

# Predicted activities for subject of average age; other factors set to lowest (reference) level
```{r}
# Status for given mean age and reference state for all categorical variables
ave.tot <-exp(coef(f.tot)[1]) # total cell numbers
ave.status <- exp(coef.matrix[,1]) # almost all cells are in dead category

# Check: proportions should sum to ~1
sum(ave.status)

# What happens when you take long showers?
l3.tot <- exp(coef(f.tot)[1])*exp(coef(f.tot)["show.l3"])

cat("Long showers reduce total bacteria by", round((ave.tot - l3.tot)/ave.tot*100, 1), "%")

# Long showers increases proportion of dormant cells and decreases dead cells by a lot! 
l3.effect <-rbind(ave.status, exp(coef.matrix[,1])*exp(coef.matrix[,"show.l3"]))
l3.diff <- l3.effect[2,] - l3.effect[1,] 
l3.effect <- rbind(l3.effect, l3.diff)
rownames(l3.effect) <- c("ave.status", "l3.effect", "diff")
l3.effect

# Eczema increases total bacteria by a lot!
ecz.tot <- exp(coef(f.tot)[1])*exp(coef(f.tot)["ecze"])
cat("Eczema increases total bacteria by", round((ecz.tot - ave.tot)/ave.tot*100, 1), "%")
cat("Eczema increases total bacteria", round(ecz.tot/ave.tot, 1), "fold")

# Eczema increases proprition of dead cells by about 16%
ecz.effect <-rbind(ave.status, exp(coef.matrix[,1])*exp(coef.matrix[,"ecze"]))
ecz.diff <- ecz.effect[2,] - ecz.effect[1,]  
ecz.effect <- rbind(ecz.effect, ecz.diff)
rownames(ecz.effect) <- c("ave.status", "ecz.effect", "diff")
ecz.effect

#Residual correlation
pairs(data.frame(active = resid(f.act), dormant = resid(f.dor), dead = resid(f.ded)))
cor <- corr.test(data.frame(active = resid(f.act), dormant = resid(f.dor), 
                dead = resid(f.ded)), method = "pearson", adjust = "none")
print(cor, digits = 4)

## The strong residual correlations among all metabolic classes suggests we are missing
## major cues the cells are using to determine activity. There are a few things this 
## could be attributed to:
## 1) Dormancy transitions can be stochastic (i.e. bet hedging). Does this make sense?
## 2) Unmeasured variables (host health, genetics, etc); Turnbaugh et al. 2007
## 3) Perhaps reliability of survey (e.g., accuracy of reporting)
## 4) Host-micobe interactions, specifically immunity
```

### Fit the rest of the SEM
```{r}
# Set up list for results
pred.fits<-list()

# Fit models
dim(X)
for(i in 1:12)pred.fits[[i]]<-glm(X[,i+2]~age,data=X,family=binomial)

# Print fits
for(i in 1:12)print(summary(pred.fits[[i]]))

# Which ones have a significant age effect?
which(sapply(pred.fits,function(x)summary(x)$coef[2,"Pr(>|z|)"])<0.052)
colnames(X)[which(sapply(pred.fits,function(x)summary(x)$coef[2,"Pr(>|z|)"])<0.052)+2]

## Look at what we found
summary(pred.fits[[6]])# older people more likely to drink alcohol
summary(pred.fits[[10]])# older people less likely to take long showers

# Shower length 15<x<30 calculate Tjur's R^2 (Don?: I think this is just l3, i.e., > 15 mins)
# I don't really know what this value tells us. 
mean(predict(pred.fits[[10]],type='response')[which(show.l3==1)])-
mean(predict(pred.fits[[10]],type='response')[which(show.l3==0)])

# Test structure
fits<-c(list(f.act,f.ded,f.dor,f.tot),pred.fits)
sapply(fits,function(x)summary(x)$coefficients)
names(fits)<-names(ages.2)[3:18]
# Lots of output. How do we interpret? Compare to line 67 above (running SEM first time)

# Set up adjacency matrix. 
# Allows us to find pairs of nodes not connected directly by an arrow,
# and therefore predicted by the model structure to be conditionally indepdentent 
# (i.e. independent after accounting for all the arrows pointing at each)
A <- matrix(0,17,16)
rownames(A) <- names(ages.2)[c(2,7:18,6,3:5)]
colnames(A) <- names(ages.2)[c(7:18,6,3:5)]

# Use fits list to fill in the adjacency matrix
for(i in 1:length(fits))A[names(coef(fits[[i]])[-1]),names(fits)[i]]<-1

# Add effect of total
A["total",c("active","dead","dormant")]<-1
print(A)

# Find missing links between our target variables 
miss.links<-NULL
for(i in 13:16)for(j in 1:i)if(A[j,i]==0)miss.links<-cbind(miss.links,
          c(rownames(A)[j],colnames(A)[i]))
print(miss.links)

# Test for significant correlations among them. 
# Remember, our initial model said these should equal zero
pvals<-rep(0,ncol(miss.links))
for(i in 1:3)pvals[i]<-cor.test(resid(fits[miss.links[2,i]][[1]]),resid(fits[miss.links[1,i]][[1]]))$p.value
data.frame(t(miss.links),pvals)

# All three are significant. Let's look at them
resid.cors<-list()
for(i in 1:3)resid.cors[[i]]<-cor.test(resid(fits[miss.links[2,i]][[1]]),resid(fits[miss.links[1,i]][[1]]))
for(i in 1:3)names(resid.cors)[i]<-paste(miss.links[1,i],miss.links[2,i],sep='-')
 #use fisher's combined test as a test of the structure of our model
1-pchisq(-2*sum(log(pvals[1:3])),2*3) #it fails...we know why.

# If we add those correlations to the model, we have no more degrees of freedom to test the 
# structure of the model. Our conclusion is that although we have found some
# mechanisms, we know now that we have we have missed some mechanisms responible
# for determining how cells are partitioned across the metabolic classes. 
```

