---
  title: "Skin Metabolic Activity"
author: "Don Schoolmaster"
date: "`r format(Sys.time(), '%d %B, %Y')`"
header-includes:
  - \usepackage{array}
output: pdf_document
geometry: margin=2.54cm
---

# Set working directory
```{r}
rm(list=ls())
getwd()
setwd("~/GitHub/skin")
```

# Read in data 
```{r}
# ages.raw <- read.table("data/skin.age.txt", sep = "\t", header = TRUE)
ages.raw.count <- read.table("data/skin.age.counts.txt", sep = "\t", header = T)

#remove subject "G1"
ages <- ages.raw.count[ ! ages.raw.count$subject == "G1", ]
```

# Create dummy variables for multinomial variables
```{r}
#shower frequency
show.f2 <- ifelse(ages$show.freq==2,1,0)
show.f3 <- ifelse(ages$show.freq==3,1,0)

#shower length
show.l2 <- ifelse(ages$show.leng==2,1,0)
show.l3 <- ifelse(ages$show.leng==3,1,0)

#last shower
show.last2 <- ifelse(ages$last.show==2,1,0)
show.last3 <- ifelse(ages$last.show==3,1,0)

#create new data.frame with all goodies.
ages.2 <- data.frame(ages[,c(1:6,11:16)], show.f2,show.f3,show.l2, 
                     show.l3,show.last2,show.last3)

#set up a design matrix that we might use later
X <- data.frame(ages.2[,-c(1,3:5)])

#center age for later interpretation
mean(ages$age)
X$age <- X$age-mean(X$age)
```

# Look at raw correlations
```{r}
cor(ages.2[,c(3,4,5)]/ages.2$total)
# this upholds idea that cells are usually in an active or dormant state
# and that dormant cells are more likely to die

# JTL makes correlation figures here
```

# Perform SEM
*** Getting non-integer warnings ***
```{r}
#Active
summary(f.act<-glm.nb(ages$active~.+offset(log(X$total)),data=X[,-2]))

#Dormant
summary(f.dor<-glm.nb(ages$dormant~.+offset(log(X$total)),data=X[,-2]))

#Dead
summary(f.ded<-glm.nb(ages$dead~.+offset(log(X$total)),data=X[,-2]))

#Total
summary(f.tot<-glm.nb(ages$total~.,data=X[,-2],maxit=1000))
```

# Predicted proportions in metabolic categories for "average" individual
```{r}
# mean age and reference state for all categorical variables
(exp(c(coef(f.act)[1],coef(f.dor)[1],coef(f.ded)[1])))

# Check: proportions should sum to ~1
sum((exp(c(coef(f.act)[1],coef(f.dor)[1],coef(f.ded)[1]))))
```

# Which variables are significantly affect metabolic classes?
*** how do we interpret e.g., show.l2 vs. show.l3? ***
```{r}
#Active
act.n <- rownames(summary(f.act)$coef)[which(summary(f.act)$coef[,4]<0.05)] 
act.p <- summary(f.act)$coef[which(summary(f.act)$coef[,4]<0.05)] 
t(data.frame(act.n, act.p))

#Dormant
dor.n <- rownames(summary(f.dor)$coef)[which(summary(f.dor)$coef[,4]<0.05)]
dor.p <-summary(f.dor)$coef[which(summary(f.dor)$coef[,4]<0.05)]
t(data.frame(dor.n,dor.p))

#Dead
ded.n <- rownames(summary(f.ded)$coef)[which(summary(f.ded)$coef[,4]<0.05)] 
ded.p <-summary(f.ded)$coef[which(summary(f.ded)$coef[,4]<0.05)]
t(data.frame(ded.n,ded.p))

#Total
tot.n <- rownames(summary(f.tot)$coef)[which(summary(f.tot)$coef[,4]<0.05)] 
tot.p <- summary(f.tot)$coef[which(summary(f.tot)$coef[,4]<0.05)] # parms
t(data.frame(tot.n,tot.p))

# age and shower length both reduce cells and affect distribution of cells
# eczema increases total but does not affect the distribution across metabolic classes
```

# Identify potential dormancy cues: look at coefs of factors that are of one sign for active and dead and the opposite sign for dormancy
```{r}
# Make matrix of coefficiencts
coef.matrix <- rbind(coef(f.act),coef(f.dor),coef(f.ded))
rownames(coef.matrix)<-c('active','dormant','dead')

# Consider variables that were significant above
coef.matrix[,"age"] #age is dormancy inducing mechanism (and reduces total cell number)
coef.matrix[,"show.f3"] #show frequncy is a dormancy reducing mechanism
coef.matrix[,"show.l3"] #show length is not a dormancy mechanism, but 
#works to preferentially remove dead cells--this interp is supported by the 
#significant effects on total cell count see detail below

# ***What about ecze, show.l2???***

# Notice that shower freq and last show should be on average inversely related and the 
# coefs show that they tell the same story shower freq tends to reduce dormancy
# so, the longer its been since showering the higher proportion is in the dormant category
```

# look at the distibution of cells into each metabolic category of a subject of average age
# and all other factors set to lowest (reference) level
# ***Not sure I understand what this does for us***
```{r}
exp(coef(f.tot)[1]) # total cell numbers
exp(coef.matrix[,1]) # almost all cells are in dead category
```

#remember there is some error which is why the proportions don't sum to exactly 1

#look at the effect of long showers, it reduces the total number of cells A LOT, disproportionately at the expence of dead ones
exp(coef(f.tot)[1])*exp(coef(f.tot)["show.l3"])
exp(coef.matrix[,1])*exp(coef.matrix[,"show.l3"])
#thus result of long showers is fewer total cells and a smaller proportion of them
#in the dead category

#eczema has a large effect on the total number of cells, but each metabolic category seems 
#to be affected proportionaltely
exp(coef(f.tot)[1]) #total cell numbers
exp(coef.matrix[,1]) #almost all cells are in dead category

exp(coef(f.tot)[1])*exp(coef(f.tot)["ecze"])
exp(coef.matrix[,1])*exp(coef.matrix[,"ecze"])

#Residual correlation
pairs(data.frame(active=resid(f.act),dormant=resid(f.dor),dead=resid(f.ded)))
cor(data.frame(active=resid(f.act),dormant=resid(f.dor),dead=resid(f.ded)))
##The strong residual correlations suggests that we are missing some
##major cues the cells are using to determine activity

.......................


####fit the rest of the SEM
#set up list for results
pred.fits<-list()
#fit models
dim(X)
for(i in 1:12)pred.fits[[i]]<-glm(X[,i+2]~age,data=X,family=binomial)
#print fits
for(i in 1:12)print(summary(pred.fits[[i]]))
#which ones have sig age effect?
which(sapply(pred.fits,function(x)summary(x)$coef[2,"Pr(>|z|)"])<0.052)
colnames(X)[which(sapply(pred.fits,function(x)summary(x)$coef[2,"Pr(>|z|)"])<0.052)+2]
##look at what we found
summary(pred.fits[[6]])#older people are more likely to drink alcohol
summary(pred.fits[[10]])#older people are less likely to take long showers
#for shower length 15<x<30 calculate Tjur's R^2
mean(predict(pred.fits[[10]],type='response')[which(show.l3==1)])-
  mean(predict(pred.fits[[10]],type='response')[which(show.l3==0)])

#test structure
fits<-c(list(f.act,f.ded,f.dor,f.tot),pred.fits)
sapply(fits,function(x)summary(x)$coefficients)
names(fits)<-names(ages.2)[3:18]
#set up adjacency matrix. This will allow us to find all the pairs of nodes 
#that are no connected directly by an arrow and therefore predicted by the model structure 
#to be conditionally indepdentant (i.e. indep. after accounting for all the arrows pointing at each)
A<-matrix(0,17,16)
rownames(A)<-names(ages.2)[c(2,7:18,6,3:5)]
colnames(A)<-names(ages.2)[c(7:18,6,3:5)]

#use fits list to fill in the adjacency matrix
for(i in 1:length(fits))A[names(coef(fits[[i]])[-1]),names(fits)[i]]<-1
#add effect of total
A["total",c("active","dead","dormant")]<-1
print(A)
#find missing links between our target variables 
miss.links<-NULL
for(i in 13:16)for(j in 1:i)if(A[j,i]==0)miss.links<-cbind(miss.links,
                                                             c(rownames(A)[j],colnames(A)[i]))
print(miss.links)
#test for significant correlations among them. Remember, our initial model said these should equal zero
pvals<-rep(0,ncol(miss.links))
for(i in 1:3)pvals[i]<-cor.test(resid(fits[miss.links[2,i]][[1]]),resid(fits[miss.links[1,i]][[1]]))$p.value
data.frame(t(miss.links),pvals)
#all three are significant. let's look at them
resid.cors<-list()
for(i in 1:3)resid.cors[[i]]<-cor.test(resid(fits[miss.links[2,i]][[1]]),resid(fits[miss.links[1,i]][[1]]))
for(i in 1:3)names(resid.cors)[i]<-paste(miss.links[1,i],miss.links[2,i],sep='-')
 #use fisher's combined test as a test of the structure of our model
1-pchisq(-2*sum(log(pvals[1:3])),2*3) #it fails...we know why.

#if we add those correlations to the model, we have no more degrees of freedom to test the 
#structure of the model. Our conclusion is that although we have found some
#mechanisms, we know now that we have we have missed some mechanisms responible
#for determining how cells are partitioned across the metabolic classes. 
