---
title: "Skin Microbiome"
author: "Jay T. Lennon"
date: "`r format(Sys.time(), '%d %B, %Y')`"
header-includes:
   - \usepackage{array}
output: pdf_document
geometry: margin=2.54cm
---

## 1) SETUP
### A. Retrieve and Set Your Working Directory

```{r, results = 'hide'}
rm(list = ls())
getwd()
setwd("~/GitHub/skin")
```

### B. Load Packages 
```{r}
require("vegan")
```

## 2) ANALYSIS OF DIFFERENT HABITATS

### A. Load data
```{r}
# Load raw data
setwd("~/GitHub/skin")
hab <- read.table("data/skin.habitat.txt", sep = "\t", header = TRUE)

# Look at raw data structure
str(hab)
```

### B. Caculate total bacterial abundances
```{r}
# Area sampled (centimeters squared) per subject
area.cm2 <- 6.4516 

# Sample volume containg skin bacteria (mL)
samp.vol <- 5

# Subsample volume that went onto filer (mL)
sub.vol <- 1.5

# Inside diameter (diam) and radius (rad) of filter tower (cm)
tow.diam <- 1.64
tow.rad <- tow.diam / 2

# Area of filter
filter.area <- pi * (tow.rad)^2

# Dimensions of field (cm) with 40x objective
field.width <- 0.0215
field.height <- 0.0160
field.cm2 <- field.width * field.height
fields.filter <- filter.area / field.cm2

# Bacterial cells per filter
cell.filter <- hab$total * fields.filter

# Bacterial densities (cm^-2)
cell.cm2 <- ((cell.filter * samp.vol) / sub.vol) / area.cm2

# Merge with habitat matrix
hab <- cbind(hab, cell.cm2)

# Minimum and maximum density of cells
cat("minimum cell density =", format(min(hab$cell.cm2), scientific = TRUE), "per cm2")
cat("maximum cell density =",format(max(hab$cell.cm2), scientific = TRUE), "per cm2")

# Make table of total abundances
sem <- function(x){
  sd(x)/sqrt(length(x))
  }

table.abund <- aggregate(hab[, c("cell.cm2")], 
            by = list(hab$habitat), function(x) c(mean=mean(x), sem=sem(x)))

format(table.abund[,1:2], scientific = TRUE)

# anova.fit <- aov(total ~ habitat + Error(subject/total), data = hab)
anova.fit <- aov(total ~ habitat, data = hab)
TukeyHSD(anova.fit)
```

### C. Calculate proportions of each metabolic class
```{r}
# Percent active, dead, and dormant
act.prop <- hab$active/hab$total
dead.prop <- hab$dead/hab$total 
dorm.prop <- hab$dormant/hab$total 
inact.prop <- (hab$dead + hab$dormant) / hab$total 

# Make new dataframe with proportion metabolic class data
hab.prop <- data.frame(hab[1:2], hab$gender, act.prop, dorm.prop, dead.prop, inact.prop)
```

### D. Make a summary table for metabolic classes by habitat
```{r}
table.hab <- aggregate(hab.prop[, c("act.prop", "dead.prop", "dorm.prop", "inact.prop")], 
            by = list(hab.prop$habitat), function(x) c(mean=mean(x), sem=sem(x)))
colnames(table.hab) <- c("habitat", "active", "dead", "dormant", "inactive")

print(table.hab)

table.gend <- aggregate(hab.prop[, c("act.prop", "dead.prop", "dorm.prop", "inact.prop")], 
            by = list(hab.prop$hab.gender), function(x) c(mean=mean(x), sem=sem(x)))
colnames(table.gend) <- c("gender", "active", "dead", "dormant", "inactive")

print(table.gend)
```

### E. Make stacked bar chart for each subject and different habitats
```{r}
# Subset habitats
back <- subset(hab.prop, habitat == "back")
back.sub <- t(back[,c(6,5,4)])
colnames(back.sub) <- 1:10

arm <- subset(hab.prop, habitat == "arm")
arm.sub <- t(arm[,c(6,5,4)])
colnames(arm.sub) <- 1:10

knee <- subset(hab.prop, habitat == "knee")
knee.sub <- t(knee[,c(6,5,4)])
colnames(knee.sub) <- 1:10

# Plot
png(filename="~/GitHub/skin/figures/skin.habitats.png",
    width = 1200, height = 1200, res = 96*2)

par(mfrow = c(1,1), mar = c(1,7,2,5.5))
bar.layout <- layout(rbind(1, 2, 3, 4), height = c(4, 4, 5, 1)) 
#layout.show(bar.layout)

barplot.back <- barplot(as.matrix(back.sub),
        ylim = c(0, 1), las = 1,
        beside = FALSE, yaxt = "n",
        cex.axis = 1.5,
        axisnames = FALSE,
        col = c("chartreuse3", "darkblue", "red"), plot = TRUE)
axis(side=2, lwd.ticks = 2, cex.axis = 1.5, las = 1, labels = c("0.0", "0.5",
              "1.0"), at = c(0, 0.5, 1.0))
mtext("Back", side = 4, las = 1, cex = 1.5, adj = 0, line = -1)

par(mar = c(1, 7, 1, 5.5))

barplot.arm <- barplot(as.matrix(arm.sub),
        ylim = c(0, 1), las = 1,
        beside = FALSE, yaxt = "n",
        cex.axis = 1.5,
        axisnames = FALSE,
        col = c("chartreuse3", "darkblue", "red"), plot = TRUE)
axis(side=2, lwd.ticks = 2, cex.axis = 1.5, las = 1, labels = c("0.0", "0.5",
              "1.0"), at = c(0, 0.5, 1.0))
mtext("Proportion", side = 2, line = 4, cex = 2)
mtext("Arm", side = 4, las = 1, cex = 1.5, adj = 0, line = -1)

par(mar = c(5, 7, 1, 5.5))
barplot.knee <- barplot(as.matrix(knee.sub),
        ylim = c(0, 1), las = 1,
        beside = FALSE, yaxt = "n",
        cex.axis = 1.5,
        xlab = "Subjects", cex.lab = 2, cex.names = 1.5,
        axisnames = TRUE, 
        col = c("chartreuse3", "darkblue", "red"), plot = TRUE)
axis(side=2, lwd.ticks = 2, cex.axis = 1.5, las = 1, labels = c("0.0", "0.5",
              "1.0"), at = c(0, 0.5, 1.0))
mtext("Knee", side = 4, las = 1, cex = 1.5, adj = 0, line = -1)

par(mar=c(0, 9, 0, 0))
plot.new()

legend(0, 1, c("Active","Dormant","Dead"), pt.lwd = 2, col = "black", 
      pt.bg = c("chartreuse3", "darkblue", "red"), pch = c(22, 22, 22), bty='n',
      ncol = 3, cex = 2, pt.cex = 5, xjust = 0)

dev.off() # this writes plot to folder
graphics.off() # shuts down open devices
```

### F. Test for differences among habitats using PERMANOVA
```{r}
# Subset activity data
hab.prop.cell <- hab.prop[,4:6]
hab.prop.act.inact <- hab.prop[,c(4,7)]
hab.type <- hab.prop$habitat
hab.gend <- hab.prop$hab.gender
subjects <- hab.prop$subject

# Run PERMANOVA with adonis function blocking by subject
adonis(hab.prop.cell ~ hab.type, method = "bray", permutations = 9999, 
       strata = subjects) 

# Run PERMANOVA with adonis function blocking by subject with gender
adonis(hab.prop.cell ~ hab.type + hab.gend + hab.type * hab.gend, 
       method = "bray", permutations = 9999, strata = subjects) 

# Run PERMANOVA with adonis function on active vs. inactive blocking by subject
adonis(hab.prop.act.inact ~ hab.type, method = "bray", permutations = 9999, 
       strata = subjects) 

# Run ANOVA on ratio of active to inactive (dead + dormant) to test for habitat
act.inact.ratio <- hab.prop.act.inact$act.prop / hab.prop.act.inact$inact.prop
anova.fit2 <- aov(act.inact.ratio ~ habitat, data = hab)
summary(anova.fit2)
TukeyHSD(anova.fit2)
```