---
title: "Skin Microbiome"
author: "Jay T. Lennon"
date: "`r format(Sys.time(), '%d %B, %Y')`"
header-includes:
   - \usepackage{array}
output: pdf_document
geometry: margin=2.54cm
---

## 1) Retrieve and Set Your Working Directory

```{r, results = 'hide'}
rm(list = ls())
getwd()
setwd("~/GitHub/skin")
```

### 2) E. coli control
```{r}
ec.kill <- c(17, 685, 3700) # active, dormant, dead
ec.kill.rel <- ec.kill / sum(ec.kill) * 100

ec.live <- c(2256, 455, 287)  # active, dormant, dead
ec.live.rel <- ec.live / sum(ec.live) * 100

## 2) Load data
```{r}
# Load raw data
ctc <- read.table("data/skin.isolate.CTC.txt", sep = "\t", header = TRUE)

# Look at raw data structure
str(ctc)
```

## 3) Subsetting Data
```{r}
# treatment info
treat <- cbind(as.data.frame(ctc$Strain), ctc$Rep, ctc$Treat)

# cells with no CTC
minus <- ctc[ctc$Treat == "minus",]

# killed cells with CTC
kill <- ctc[ctc$Treat == "kill",]

# cells incubated with CTC
plus <- ctc[ctc$Treat == "plus",]
```

## 4) Correcting and making new dataframe
```{r}
# Correct dtat by controls

# kill-corrected
kill.corr <- kill$Fluor - minus$Fluor

# plus-corrected
plus.corr <- plus$Fluor - minus$Fluor

# Create strains vector
strains <- c("K1", "K1", "K1", "K2", "K2", "K2", "K3", "K3", "K3",
             "F1", "F1", "F1", "F2", "F2", "F2", "B1", "B1", "B1", "K4",
             "K4", "K4", "K5", "K5","K5", "K6", "K6","K6","EC","EC", "EC")

# Merge into new data frame
data.corr <- as.data.frame(cbind(strains, kill.corr, plus.corr))
data.corr$kill.corr <- as.numeric(as.character(data.corr$kill.corr))
data.corr$plus.corr <- as.numeric(as.character(data.corr$plus.corr))
as.data.frame(data.corr)
```

## 5) Multiple paired t-test
```{r}
# Create output matrix for summary statistics
test.out <- matrix(data =NA, length(levels(data.corr$strains)), 3)
row.names(test.out) <- c("K1", "K2", "K3", "F1", "F2", "B1", "K4", "K5", "K6", "EC")
colnames(test.out) <- c("t.value", "df", "p.value")

# For loop to run t-tests
for(i in levels(data.corr$strains)) {
  tmp = data.corr[data.corr$strains == i,]
  test = t.test(tmp$kill.corr, tmp$plus.corr, paired = TRUE)
  test.out[i, "t.value"] <- test$statistic
  test.out[i, "df"] <- test$parameter
  test.out[i, "p.value"] <- test$p.value
  }

# Benjamini & Hochberg correction for multiple comparisons
p.adjust <- p.adjust(test.out[,3], method = "BH", n = length(test.out[,3]))

# Final output matrix with corrected p-values
test.out.final <- cbind(test.out, p.adjust) 
# For loop to run t-tests
```
  
## 6) Means and figure
```{r}
summary.out <- matrix(data = NA, length(unique(strains)), 4)
row.names(summary.out) <- c("K1", "K2", "K3", "F1", "F2", "B1", "K4", "K5", "K6", "EC")
colnames(summary.out) <- c("mean.dead", "sem.dead", "mean.live", "sem.live")
sem <- function(x) sqrt(var(x)/length(x))

for(i in levels(data.corr$strains)) {
  tmp = data.corr[data.corr$strains == i,]  
      test = c(mean(tmp$kill.corr), sem(tmp$kill.corr), mean(tmp$plus.corr), sem(tmp$plus.corr))
  summary.out[i, "mean.dead"] <- test[1]
  summary.out[i, "sem.dead"] <- test[2]
  summary.out[i, "mean.live"] <- test[3]
  summary.out[i, "sem.live"] <- test[4]
  }

# Fold differences
fold.out <- summary.out[,3] / summary.out[,1]
mean.fold.out <- mean(fold.out)
sem.fold.out <- sem(fold.out)

x.dead <- c(0, 25, 50, 75, 100, 125, 150, 175, 200, 225)
x.live <- x.dead 

plot.new()
plot(x.dead, summary.out[,1], type = "p", pch = 22, bg = "white", col = "black", 
     las = 1, xlim = c(-20, 250), ylim = c(-500, 60000), cex = 2, yaxt = "n", 
     cex.lab = 2, cex.axis = 1.5, ylab = "", xlab = "", yaxt = "n", xaxt = "n",
     lwd = 2,
     panel.first={
       arrows(x0 = x.dead, y0 = summary.out[,1], y1 = summary.out[,1] - summary.out[,2], angle = 90, 
       length = 0.05, lwd = 2)
       arrows(x0 = x.dead, y0 = summary.out[,1], y1 = summary.out[,1] + summary.out[,2], angle = 90, 
       length = 0.05, lwd = 2)
})  
     


points(x.live, summary.out[,3], type = "p", pch = 22, bg = "black", 
       col = "black", cex = 2,

  panel.first={
       arrows(x0 = x.dead, y0 = summary.out[,3], y1 = summary.out[,3] - summary.out[,4], angle = 90, 
       length = 0.05, lwd = 2)
       arrows(x0 = x.dead, y0 = summary.out[,3], y1 = summary.out[,3] + summary.out[,4], angle = 90, 
       length = 0.05, lwd = 2)
})  

# Major Axes
axis(side = 2, lwd.ticks = 2, cex.axis = 1.5, las = 1,
    labels = c(0, 25000, 50000), at = c(0, 25000, 60000))
 
axis(side = 4, lwd.ticks = 2, cex.axis = 1.5, las = 1,
    at=c(0, 25000, 50000), labels = F)
 
axis(side = 1, lwd.ticks = 2, cex.axis = 1, las = 1,
    labels = c("K1", "K2", "K3", "F1", "F2", "B1", "K4", "K5", "K6", "EC"), 
    at = c(0, 25, 50, 75, 100, 125, 150, 175, 200, 225))

axis(side = 3, lwd.ticks = 2, cex.axis = 1.5, las = 1,
    at = c(0, 25, 50, 75, 100, 125, 150, 175, 200, 225), labels = F)

box(lwd = 2)
```

