---
title: "Skin Metabolic Activity"
author: "Donald R. Schoolmaster Jr. and Jay T. Lennon"
date: "`r format(Sys.time(), '%d %B, %Y')`"
header-includes:
  - \usepackage{array}
output: pdf_document
geometry: margin=2.54cm
---

# Set working directory and load packages
```{r}
rm(list=ls())
getwd()
setwd("~/GitHub/skin")
require("psych")
require("rgl")
require("scatterplot3d")
require("ggtern")
```

# Read in data 
```{r}
 # Read in raw data
ages.raw.count <- read.table("data/skin.age.counts.txt", sep = "\t", header = T)

# Remove subject "G1"
ages <- ages.raw.count[ ! ages.raw.count$subject == "G1", ]

# Stadnardize active, dormant, and dead by total abundance
ages.st <- data.frame(ages[,1:2], ages[,3:5]/ages[,6], ages[,6:16])
```

# Look at correlations
```{r}
cor <- corr.test(ages.st[,3:5], method = "pearson", adjust = "none")
print(cor, digits = 3)

dorm.act.p <- ifelse(cor$p[3] == 0, '< 0.0001', round(cor$p[3], digits = 3))
dorm.act.r <- round(cor$r[3], digits = 2)

dead.act.p <- ifelse(cor$p[2] == 0, '< 0.0001', round(cor$p[2], digits = 3))
dead.act.r <- round(cor$r[2], digits = 2)

dead.dorm.p <- ifelse(cor$p[6] == 0, '< 0.0001', round(cor$p[6], digits = 3))
dead.dorm.r <- round(cor$r[6], digits = 2)

# this upholds idea that cells are usually in an active or dormant state
# and that dormant cells are more likely to die
```

# Plot correlations (standardized by total abundance)
```{r}
png(filename="figures/skin.correlations.png",
    width = 1200, height = 500, res = 96*2)

layout(matrix(c(1,2,3), 1, 3, byrow = TRUE))

# Dormant vs. Active
plot(ages.st$active, ages.st$dormant, pch = 22, bg = "white", col = "black", 
     xlab = NA, ylab = NA, xlim = c(0, 1), ylim = c(0,1), cex = 2, lwd = 2,
     yaxt = "n", xaxt = "n", las = 2)
box(lwd = 2)

axis(side = 2, lwd.ticks = 2, cex.axis = 1.0, las = 1,
    labels = c("0.0", "0.5", "1.0"), at = c(0, 0.5, 1.0))
 
axis(side = 4, lwd.ticks = 2, cex.axis = 1.0, las = 1,
    at=c(0, 0.5, 1), labels = F)
 
axis(side = 1, lwd.ticks = 2, cex.axis = 1.0, las = 1,
    labels = c("0.0", "0.5", "1.0"), at = c(0, 0.5, 1.0))

axis(side = 3, lwd.ticks = 2, cex.axis = 1.0, las = 1,
    at=c(0, 0.5, 1), labels = F)

mtext("Dormant", side = 2, cex = 0.75, 
      line = 2.5, adj = 0.5)

mtext("Active", side = 1, cex = 0.75, 
      line = 3, adj = 0.5)
      
dorm.act.r.label <- bquote(italic(r) == .(format(dorm.act.r)))
text(x = 0.60, y = 0.95, labels = dorm.act.r.label, adj = 0)

dorm.act.p.label <- bquote(italic(P) == .(format(dorm.act.p)))
text(x = 0.60, y = 0.85, labels = dorm.act.p.label, adj = 0)

# Dead vs. Active
plot(ages.st$active, ages.st$dead, pch = 22, bg = "white", col = "black", 
     xlab = NA, ylab = NA, xlim = c(0, 1), ylim = c(0,1), cex = 2, lwd = 2,
     yaxt = "n", xaxt = "n", las = 2)     
box(lwd = 2)

axis(side = 2, lwd.ticks = 2, cex.axis = 1.0, las = 1,
    labels = c("0.0", "0.5", "1.0"), at = c(0, 0.5, 1.0))
 
axis(side = 4, lwd.ticks = 2, cex.axis = 1.0, las = 1,
    at=c(0, 0.5, 1), labels = F)
 
axis(side = 1, lwd.ticks = 2, cex.axis = 1.0, las = 1,
    labels = c("0.0", "0.5", "1.0"), at = c(0, 0.5, 1.0))

axis(side = 3, lwd.ticks = 2, cex.axis = 1.0, las = 1,
    at=c(0, 0.5, 1), labels = F)

mtext("Dead", side = 2, cex = 0.75, 
      line = 2.5, adj = 0.5)

mtext("Active", side = 1, cex = 0.75, 
      line = 3, adj = 0.5)

dead.act.r.label <- bquote(italic(r) == .(format(dead.act.r)))
text(x = 0.6, y = 0.95, labels = dead.act.r.label, adj = 0)

dead.act.p.label <- bquote(italic(P) == .(format(dead.act.p)))
text(x = 0.6, y = 0.85, labels = dead.act.p.label, adj = 0)

# Dead vs. Dormant
plot(ages.st$dormant, ages.st$dead, pch = 22, bg = "white", col = "black", 
     xlab = NA, ylab = NA, xlim = c(0, 1), ylim = c(0,1), cex = 2, lwd = 2,
     yaxt = "n", xaxt = "n", las = 2)     
box(lwd = 2)

axis(side = 2, lwd.ticks = 2, cex.axis = 1.0, las = 1,
    labels = c("0.0", "0.5", "1.0"), at = c(0, 0.5, 1.0))
 
axis(side = 4, lwd.ticks = 2, cex.axis = 1.0, las = 1,
    at=c(0, 0.5, 1), labels = F)
 
axis(side = 1, lwd.ticks = 2, cex.axis = 1.0, las = 1,
    labels = c("0.0", "0.5", "1.0"), at = c(0, 0.5, 1.0))

axis(side = 3, lwd.ticks = 2, cex.axis = 1.0, las = 1,
    at=c(0, 0.5, 1), labels = F)

mtext("Dead", side = 2, cex = 0.75, 
      line = 2.5, adj = 0.5)

mtext("Dormant", side = 1, cex = 0.75, 
      line = 3, adj = 0.5)

dead.dorm.r.label <- bquote(italic(r) == .(format(dead.dorm.r)))
text(x = 0.45, y = 0.95, labels = dead.dorm.r.label, adj = 0)

dead.dorm.p.label <- bquote(italic(P) == .(format(dead.dorm.p)))
text(x = 0.45, y = 0.85, labels = dead.dorm.p.label, adj = 0)


# Close Plot Device
dev.off()
graphics.off()
```

# Some 3-D plotting options
```{r}
active <-ages.st$active
dormant <- ages.st$dormant
dead <- ages.st$dead

scatterplot3d(active, dormant, dead, pch = 22, color = "black")
plot3d(active, dormant, dead, col="black", size=5)
# ggtern(data=ages,aes(active,dormant,dead)) + 
#   geom_point(fill="red",shape = 21, size = 4) + 
#   theme_tern_bw()
```


# Plot correlations (standardized by total abundance)
```{r}
png(filename="figures/skin.correlations.png",
    width = 1200, height = 500, res = 96*2)

layout(matrix(c(1,2,3), 1, 3, byrow = TRUE))

# Dormant vs. Active
plot(ages$active, ages$dormant, pch = 22, bg = "white",
     col = "black", log = "xy", xlab = NA, ylab = NA, 
     xlim = c(5000, 10000000), ylim = c(50000,10000000), yaxt = "n", xaxt = "n",
     cex = 2, lwd = 2)
box(lwd = 2)

axis(side = 2, lwd.ticks = 2, cex.axis = 1, las = 1,
     labels = expression(10^4, 10^5, 10^6, 10^7),
     at = c(10^4, 10^5, 10^6, 10^7))
 
axis(side = 4, lwd.ticks = 2, cex.axis = 1, las = 1,
    at=c(10^4, 10^5, 10^6, 10^7), labels = F)
 
axis(side = 1, lwd.ticks = 2, cex.axis = 1, las = 1,
     labels = expression(10^4, 10^5, 10^6, 10^7),
     at = c(10^4, 10^5, 10^6, 10^7))

axis(side = 3, lwd.ticks = 2, cex.axis = 1, las = 1,
    at = c(10^4, 10^5, 10^6, 10^7), labels = F)

mtext(expression('Dormant (cells cm'^-2*')'), side = 2, cex = 0.75, 
      line = 2.5, adj = 0.5)

mtext(expression('Active (cells cm'^-2*')'), side = 1, cex = 0.75, 
      line = 3, adj = 0.5)
      
dorm.act.r.label <- bquote(italic(r) == .(format(dorm.act.r)))
text(x = 5*10^6, y = 2*10^5, labels = dorm.act.r.label)

dorm.act.p.label <- bquote(italic(P) == .(format(dorm.act.p)))
text(x = 5*10^6, y = 1*10^5, labels = dorm.act.p.label)

# Dead vs. Active
plot(ages$active, ages$dead, pch = 22, bg = "white",
     col = "black", log = "xy", xlab = NA, ylab = NA, 
     xlim = c(5000, 10000000), ylim = c(50000,100000000), 
     yaxt = "n", xaxt = "n", cex = 2, lwd = 2)
     
box(lwd = 2)

axis(side = 2, lwd.ticks = 2, cex.axis = 1, las = 1,
     labels = expression(10^4, 10^5, 10^6, 10^7, 10^8),
     at = c(10^4, 10^5, 10^6, 10^7,10^8))
 
axis(side = 4, lwd.ticks = 2, cex.axis = 1, las = 1,
    at=c(10^4, 10^5, 10^6, 10^7, 10^8), labels = F)
 
axis(side = 1, lwd.ticks = 2, cex.axis = 1, las = 1,
     labels = expression(10^4, 10^5, 10^6, 10^7),
     at = c(10^4, 10^5, 10^6, 10^7))

axis(side = 3, lwd.ticks = 2, cex.axis = 1, las = 1,
    at = c(10^4, 10^5, 10^6, 10^7), labels = F)

mtext(expression('Dead (cells cm'^-2*')'), side = 2, cex = 0.75, 
      line = 2.5, adj = 0.5)

mtext(expression('Active (cells cm'^-2*')'), side = 1, cex = 0.75, 
      line = 3, adj = 0.5)

dead.act.r.label <- bquote(italic(r) == .(format(dead.act.r)))
text(x = 5*10^6, y = 2*10^5, labels = dead.act.r.label)

dead.act.p.label <- bquote(italic(P) == .(format(dead.act.p)))
text(x = 5*10^6, y = 1*10^5, labels = dead.act.p.label)

# Dead vs. Dormant
plot(ages$dormant, ages$dead, pch = 22, bg = "white",
     col = "black", log = "xy", xlab = NA, ylab = NA, 
     xlim = c(100000, 10000000), ylim = c(50000,100000000), 
     yaxt = "n", xaxt = "n", cex = 2, lwd = 2)
     
box(lwd = 2)

axis(side = 2, lwd.ticks = 2, cex.axis = 1, las = 1,
     labels = expression(10^4, 10^5, 10^6, 10^7, 10^8),
     at = c(10^4, 10^5, 10^6, 10^7,10^8))
 
axis(side = 4, lwd.ticks = 2, cex.axis = 1, las = 1,
    at=c(10^4, 10^5, 10^6, 10^7, 10^8), labels = F)
 
axis(side = 1, lwd.ticks = 2, cex.axis = 1, las = 1,
     labels = expression(10^4, 10^5, 10^6, 10^7),
     at = c(10^4, 10^5, 10^6, 10^7))

axis(side = 3, lwd.ticks = 2, cex.axis = 1, las = 1,
    at = c(10^4, 10^5, 10^6, 10^7), labels = F)

mtext(expression('Dormant (cells cm'^-2*')'), side = 1, cex = 0.75, 
      line = 2.5, adj = 0.5)

mtext(expression('Dead (cells cm'^-2*')'), side = 2, cex = 0.75, 
      line = 3, adj = 0.5)

dead.dorm.r.label <- bquote(italic(r) == .(format(dead.dorm.r)))
text(x = 5*10^6, y = 2*10^5, labels = dead.dorm.r.label)

dead.dorm.p.label <- bquote(italic(P) == .(format(dead.dorm.p)))
text(x = 5*10^6, y = 1*10^5, labels = dead.dorm.p.label)

# Close Plot Device
dev.off()
graphics.off()
